%!TEX root = ../main.tex

\section*{Data Collection}\label{sec:data}

% Data collection for study two involves five steps. First, a timeline of events is created \citep{Maguire2009}. From that timeline, we select forums that facilitate a discourse \citep{Latour2005}. Then, texts from the participating communities on the discourse are gathered. The data analysis is carried out on the basis of these texts. Finally, I interview individual members of the communities to test the robustness of my fiends and ensure convergence.

This section introduces the data collection for study two. To understand the discourse on Keystone XL, and how it unfolded over time, we require a sample of texts that encompasses multiple relevant actors over an extended period of time. The first step toward building this corpus is to develop a timeline of events \citep{Maguire2009}. The basis for that timeline are newspaper articles from the sampling period of 2008-2021. These articles were obtained from five major American national news outlets
% The Wall Street Journal, The New York Times, Los Angeles Times, and The Washington Post 
through Factiva by searching for the Keywords "Keystone XL", and "KXL". Further, articles on Keystone XL over the sampling period were collected from the Oil \& Gas Journal. 
% An excerpt of the timeline can be seen in \autoref{tab:timeline}.

% \input{tables/timeline}

The timeline on Keystone XL includes multiple public forums \citep{Latour2005}. These public forums serve two functions in this study. (1) They allow for the identification of relevant participants in the discourse. 12 types of participants participate in the forums, 6 of which are specific organizations that participate repeatedly. 
% (see \autoref{tab:forums}). 
(2) The public forums are a compressed version of the discourse itself, where actors lay out their understanding of the permit application process, of Keystone XL, and of its environmental impact. The actors also respond to each others claims in this forum, laying bare the extent to which they have engaged with each others' arguments. I.e., a rebuttal testimony begins with a summary of the other side's position, and make a useful gauge of engagement and comprehension.

% \input{tables/forums}

By selecting on the participation in forums, we obtain a sample of actors that are both knowledgeable and engaged. To participate in the public forum, an organization needs to both familiarize itself with the formal process and strategize, as well as obtain and organize expert knowledge. 
% Some of the participants, such as the Sierra Club or Bold Nebraska are or become communications experts over the course of the discourse.

The final step toward creating a sample of texts is to gather the communication of the actors on Keystone XL over the time of the discourse. Most of the communities have formed and organize themselves through websites. Documents on TransCanada, Keystone XL, and the Sierra Club are also published online. The documents are downloaded by hand, or automatically through web scraping where appropriate. In some cases, the texts are automatically filtered to retain only those that discuss Keystone XL, using the keywords "Keystone XL" and "KXL". The filtered texts, and the texts from step two constitute the final sample that is imported into ATLAS.ti for the data analysis.

% Some of the forums were open to public participation. Comments by residents were also collected but were too numerous to manually sight. Topic modeling and sentiment analysis were used to identify substreams of that corpus. Samples from each substream then entered the analysis.