\documentclass{article}

\usepackage[natbib=true, style=apa, alldates=year]{biblatex}
\usepackage[pdfauthor={Julian Barg},pdftitle={Discourse--Theory and Literature}]{hyperref}

\addbibresource{bibliography.bib}

\title{Discourse--Theory and Literature}

\author{
	Barg, Julian\\
	\texttt{jbarg.phd@ivey.ca}\\
	Ivey Business School
}

\begin{document}

	\maketitle
	\clearpage

	\input{sections/overview}

	\input{sections/translation}

	\input{sections/production_of_texts}

	\subsection*{Research Opportunity}

	We can build on Maguire \& Hardy's empirical research on discourse and extend their models by focusing on the construction of accounts. What texts and accounts become "officially sanctioned" and enter the discourse? My last note included five empirical examples that I will interpret accordingly.

	\begin{quote}
		(1) In an official meeting of the Nebraska Natural Resource Committee between the legislature, TransCanada, and the Sierra Club, the participants celebrate the latest decision to reroute Keystone XL and conduct a state-funded Environmental Impact Assessment. Senator Mark Christensen presses for Ken Winston of the Sierra Club to abide by the new status quo and cease resistance against the new route, since the demands of the Sierra Club were met \citep{NRC2011}.
	\end{quote}

	Previous exchanges between the Sierra Club and the Nebraska legislature focused on the rerouting of the Keystone XL pipeline. Senator Mark Christensen attempts to exert social pressure on the representative of the Sierra Club and hold him to his past stance as calls for the cancellation of Keystone XL are on the horizon.

	\begin{quote}
		(2) A public hearing on Keystone XL in Nebraska. A majority of participants seems concerned about the issue of pipeline safety. All questions concerning pipeline safety are rejected, since it is a matter for a national agency (PHSMA).
	\end{quote}

	In the existing institutional order, pipeline safety falls under the jurisdiction of the Pipeline and Hazardous Materials Safety Administration (PHSMA). The Nebraska Department of Environmental Quality\footnote{At least I believe they organized the meeting...} effectively enforces this segmentation of the discourse through formal rules. PHSMA holds no public hearings on the issue, effectively limiting the scope of the discourse.

	\begin{quote}
		(3) Following the President's decision to revoke the permit for Keystone XL, TransCanada files a claim with the NAFTA arbitration in the amount of \$15 billion.
	\end{quote}

	Segments of the discourse on Keystone XL are regulated through formal rules. The NAFTA arbitration institution might find that under these formal rules, the revocation of the permit by the president is not a sanctioned contribution to the discourse.

	\begin{quote}
		(4) Although indigenous groups make for some of the most persistent opponents of Keystone XL, their voice is rarely heard in the official discourse. Many of their claims revolve around treaty rights--the relevant authority is the national government.
	\end{quote}

	The discourse on treaty rights between indigenous groups goes back decades. Their claims to the lands are not taken for granted and not sanctioned to enter into the discourse. Note--this issue is much more salient with the Dakota Access Pipeline.

	\begin{quote}
		(5) Ahead of planned protests in Montana, governmental agencies create a "joint terrorism task force". A police department vows to stop the protesters "by any means".
	\end{quote}

	An additional level of conflict becomes relevant as presumably conservative leaning police forces and more liberal-minded protesters are about to clash. Terms such as "tree hugger" highlight that in the conservative eye, environmental protesters may be seen as anti-modernist and illusory--in other words, as a potentially dangerous mindset that should be suppressed. For the protesters, successful nonviolent action is a way of breaking into the discourse without having to follow the many formal rules that have been set up as barriers around it.

	\clearpage
	\section*{References}
	\printbibliography[heading=none]

\end{document}